---
title: "In-class_Ex4"
---

## **Overview**

-   performing geocoding using data downloaded from data.gov.sg

-   calibrating Geographically Weighted Poisson Regression

## **Getting Started**

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse, knitr, plotly, httr, dplyr)
```

## **Geocoding using SLA API**

Geocoding - process of aspatial description of a location e.g. address

```{r}

url <- "https://www.onemap.gov.sg/api/common/elastic/search"
csv <- read_csv("Data/Aspatial/Generalinformationofschools.csv")
postcodes <- csv$`postal_code`

found <- data.frame()
not_found <- data.frame()

for(postcode in postcodes){
  query<-list('searchVal'=postcode,'returnGeom'='Y','getAddrDetails'='Y', 'pageNum'='1')
  res <- GET(url,query=query)
  
  if((content(res)$found)!=0){
    found <- rbind(found, data.frame(content(res))[4:13])
  } else{
    not_found = data.frame(postcode)
  }
}
```

```{r}
merged = merge(csv, found, by.x= 'postal_code', by.y = 'results.POSTAL', all=TRUE)
write.csv(merged, file = "Data/Aspatial/schools.csv")
write.csv(not_found, file = "Data/Aspatial/not_found.csv")
```

### Importing geospatial data

```{r}
mpsz <- st_read(dsn = "Data/Geospatial",
                layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

### **Importing & tidying school data**

```{r}
schools <- read_csv("Data/Aspatial/Generalinformationofschools.csv")

```

Note: Zhenghua Secondary School missing. Manually update latitude and longitude for ZHENGHUA SECONDARY SCHOOL

-   Latitude: 1.389279

-   Longitude: 103.7651

```{r}

merged <- merged %>%
  mutate(
    results.LATITUDE = ifelse(school_name == "ZHENGHUA SECONDARY SCHOOL", 1.389279, results.LATITUDE),
    results.LONGITUDE = ifelse(school_name == "ZHENGHUA SECONDARY SCHOOL", 103.7651, results.LONGITUDE)
  )
```

```{r}

schools <- merged %>%
  rename(latitude = "results.LATITUDE", longitude = "results.LONGITUDE") %>%
  select(postal_code, school_name, latitude, longitude)
```

Note: schools is a tibble dataframe object class

### **Converting an aspatial data into a simple features tibble data.frame**

```{r}
schools_sf <- st_as_sf(schools,
                       coords = c("longitude", "latitude"),
                       crs =4326) %>%
  st_transform(crs = 3414)
```

### **Plotting a point simple feature layer**

```{r}

#| code-fold: true
#| code-sumary: "Show the code chunk"
tmap_mode("view")

tm_shape(schools_sf) +
  tm_dots() +
tm_view(set.zoom.limits = c(11,14))

tmap_mode("plot")
```

## **Preparing**

```{r}
mpsz$`SCHOOL_COUNT` <- lengths(
  st_intersects(
    mpsz, schools_sf
  )
)
```

Summary statistics

```{r}
summary(mpsz$SCHOOL_COUNT)
```

```{r}
business_sf <- st_read(dsn = "data/geospatial", layer = "Business") %>%
  st_transform(crs = 3414)
```

```{r}
tmap_options(check.and.fix = TRUE)
tm_shape(mpsz) +
  tm_polygons() +
tm_shape(business_sf) +
  tm_dots()
```

## Data Integration and Wrangling

Import the rds file into R environment

```{r}
flow_data <- read_rds("data/rds/flow_data_tidy.rds") 
```

```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ,
  0, flow_data$MORNING_PEAK)
flow_data$offset <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ,
  0.000001, 1)
```

```{r}
flow_data <- flow_data %>%
  filter(FlowNoIntra > 0)

```

```{r}
inter_zonal_flow <- flow_data %>%
  rename(TRIPS =MORNING_PEAK,
         DIST = dist)
```

# Model Calibration

```{r}
pacman::p_load(sp, DT, stplanr,
               performance, reshape2,
               ggpubr, units)
```

## Unconstrained SIM

```{r}
uncSIM <- glm(formula = TRIPS ~ 
                log(SCHOOL_COUNT) + 
                log(RETAIL_COUNT) +
                log(DIST),
              family = poisson(link = "log"),
              data = inter_zonal_flow,
              na.action = na.exclude)
uncSIM
```

### Goodness-of-Fit

```{r}
CalcRSquared <- function(observed,estimated){
  r <- cor(observed,estimated)
  R2 <- r^2
  R2
}
```

```{r}
CalcRSquared(uncSIM$data$TRIPS, uncSIM$fitted.values)
```

```{r}
r2_mcfadden(uncSIM)
```

## Origin constrained SIM

```{r}
orcSIM <- glm(formula = TRIPS ~
                ORIGIN_SZ +
                log(SCHOOL_COUNT) +
                log(RETAIL_COUNT) +
                log(DIST) - 1,
              family = poisson(link = "log"),
              data = inter_zonal_flow,
              na.action = na.exclude)
summary(orcSIM)
```

### Goodness-of-Fit

```{r}
CalcRSquared <- function(observed, estimated){
  r  <- cor(observed,estimated)
  R2 <- r^2
  R2
}
```

Examine how the constraints hold for destinations

```{r}
CalcRSquared(orcSIM$data$TRIPS, orcSIM$fitted.values)
```

## Destination Constrained

```{r}
decSIM <- glm(formula = TRIPS ~ 
                DESTIN_SZ + 
                log(SCHOOL_COUNT) + 
                log(DIST),
              family = poisson(link = "log"),
              data = inter_zonal_flow,
              na.action = na.exclude)
summary(decSIM)
```

```{r}
CalcRSquared(decSIM$data$TRIPS, decSIM$fitted.values)
```

## Doubly Constrained

```{r}
dbcSIM <- glm(formula = TRIPS ~
                ORIGIN_SZ +
                DESTIN_SZ +
                log(DIST),
              family = poisson(link = "log"),
              data = inter_zonal_flow,
              na.action = na.exclude)
dbcSIM
```

## Model Comparison

```{r}
model_list <- list(unconstrained=uncSIM,
                   originConstrained=orcSIM,
                   destinationConstrained=decSIM,
                   doublyConstrained=dbcSIM)

compare_performance(model_list,
                    metrics = "RMSE")
```

## Visualising Fitted

Extract the fitted values from each model

Unconstrainted Model

```{r}
df <- as.data.frame(uncSIM$fitted.values) %>%
  round(digits = 0)

inter_zonal_flow <- inter_zonal_flow %>%
  cbind(df) %>%
  rename(uncTRIPS = `uncSIM.fitted.values`)
```

Origin Constrained Model

```{r}
df <- as.data.frame(orcSIM$fitted.values) %>%
  round(digits = 0)

inter_zonal_flow <- inter_zonal_flow %>%
  cbind(df) %>%
  rename(orcTRIPS = `orcSIM.fitted.values`)
```

Destination Constrained Model

```{r}
df <- as.data.frame(decSIM$fitted.values) %>%
  round(digits = 0)

inter_zonal_flow <- inter_zonal_flow %>%
  cbind(df) %>%
  rename(decTRIPS = `decSIM.fitted.values`)
```

Doubly Constrained Model

```{r}
df <- as.data.frame(dbcSIM$fitted.values) %>%
  round(digits = 0)

inter_zonal_flow <- inter_zonal_flow %>%
  cbind(df) %>%
  rename(dbcTRIPS = `dbcSIM.fitted.values`)
```

Plotting graphs

```{r}
unc_p <- ggplot(data = inter_zonal_flow,
                aes(x = uncTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

orc_p <- ggplot(data = inter_zonal_flow,
                aes(x = orcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dec_p <- ggplot(data = inter_zonal_flow,
                aes(x = decTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dbc_p <- ggplot(data = inter_zonal_flow,
                aes(x = dbcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

ggarrange(unc_p, orc_p, dec_p, dbc_p,
          ncol = 2,
          nrow = 2)
```
