---
title: "Take-home-project"
subtitle: "Geospatial Proj"
date: "27 Nov 2023"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

## Getting Started

Lets make make sure that **sfdep**, **sf**, **tmap** and **tidyverse** packages of R are currently installed 

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse, knitr, plotly, zoo, Kendall, mapview)
```

### Importing OD data into R
Firstly we will import the Passenger volume by Origin Destination Bus Stops data set downloaded from LTA DataMall by using read_csv() of readr package.

```{r}
odbus <- read_csv("data/aspatial/origin_destination_bus_202308.csv")
glimpse(odbus)
```
```{r}
origtrip_6_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 6 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
kable(head(origtrip_6_9))
write_rds(origtrip_6_9, "data/rds/origtrip_6_9.rds")
```


```{r}
origtrip_17_20 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 17 &
           TIME_PER_HOUR <= 20) %>%
  group_by(ORIGIN_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
glimpse(origtrip_17_20)
```
```{r}
kable(head(origtrip_17_20))
write_rds(origtrip_17_20, "data/rds/origtrip_17_20.rds")
```

```{r}
origtrip_11_14 <- odbus %>%
  filter(DAY_TYPE == "WEEKENDS/HOLIDAY") %>%
  filter(TIME_PER_HOUR >= 11 &
           TIME_PER_HOUR <= 14) %>%
  group_by(ORIGIN_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
kable(head(origtrip_11_14))
write_rds(origtrip_11_14, "data/rds/origtrip_11_14.rds")
```

```{r}
origtrip_16_19 <- odbus %>%
  filter(DAY_TYPE == "WEEKENDS/HOLIDAY") %>%
  filter(TIME_PER_HOUR >= 16 &
           TIME_PER_HOUR <= 19) %>%
  group_by(ORIGIN_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
kable(head(origtrip_16_19))
write_rds(origtrip_16_19, "data/rds/origtrip_16_19.rds")
```

```{r}
origtrip_11_14 <- read_rds("data/rds/origtrip_11_14.rds")
origtrip_16_19 <- read_rds("data/rds/origtrip_16_19.rds")
origtrip_17_20 <- read_rds("data/rds/origtrip_17_20.rds")
origtrip_6_9 <- read_rds("data/rds/origtrip_6_9.rds")
glimpse(origtrip_6_9)
```

### Importing geospatial data
```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)
```
#### Glimpse the Bus Stop tibble data frame
```{r}
glimpse(busstop)
```
#### Load Map into MPSZ
```{r}
mpsz <- st_read(dsn = "data/geospatial",
                   layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```
```{r}
glimpse(mpsz)
```
## Geospatial Data Wrangling
### Combining Busstop and mpsz
```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
glimpse(busstop_mpsz)
```

```{r}
write_rds(busstop_mpsz, "data/rds/busstop_mpsz.csv")  
```

```{r}
origin_6_9 <- left_join(origtrip_6_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C) %>%
  group_by(ORIGIN_SZ) %>%
  summarise(TOT_TRIPS = sum(TRIPS))
glimpse(origin_6_9)
```

```{r}
origin_17_20 <- left_join(origtrip_17_20 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C) %>%
  group_by(ORIGIN_SZ) %>%
  summarise(TOT_TRIPS = sum(TRIPS))
```

```{r}
class(origin_17_20)
```


```{r}
origin_11_14 <- left_join(origtrip_11_14 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C) %>%
  group_by(ORIGIN_SZ) %>%
  summarise(TOT_TRIPS = sum(TRIPS))
```

```{r}
origin_16_19 <- left_join(origtrip_16_19 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C) %>%
  group_by(ORIGIN_SZ) %>%
  summarise(TOT_TRIPS = sum(TRIPS))
```

## Choropleth Visualisation
### Plotting Hexagonal Chunks





