---
title: "Hands-on_ex5"
date: "16 December 2023"
date-modified: "last-modified"
format: html
execute:
  echo: true
  warning: false
editor: visual
---

#Loading R packages

```{r}
devtools::install_github("LukeCe/spflow")
```

Next, we will load spflow and other R packages into R environment

```{r}
pacman::p_load(tmap, sf, sp, DT, stplanr,
               performance,
               ggpubr, tidyverse, httr, Matrix, spflow)
```

## Data Preparation
3 data sets are required to calaibrate a SPatial Econometric Interaction Model:
1.) A Spatial Weight
2.) A tibble data.frame cosnisiting of origins, destination, flows and distances between origins and destination
3.) Tiblee data.frame consisiting of the explanatory variables

```{r}
centroids <- suppressWarnings({
  st_point_onsurace(st_geometry(mpsz_bustop))}
  )

mpsz_nb <- list(
  "by_contiguity" = poly2nb(mpsz_busstop),
  "by_distance" = dnearneigh(centroids, 
                             d1 = 0, d2 = 5000),
  "by_knn" = knn2nb(knearneigh(centroids, 3))
)

```
