---
title: "Hands-on_ex5"
date: "16 December 2023"
date-modified: "last-modified"
format: html
execute:
  echo: true
  warning: false
editor: visual
---

#Loading R packages

```{r}
devtools::install_github("LukeCe/spflow")
```

Next, we will load spflow and other R packages into R environment

```{r}
pacman::p_load(tmap, sf, sp, DT, stplanr,
               performance,
               ggpubr, tidyverse, httr, Matrix, spflow, spdep)
```

## Data Preparation
3 data sets are required to calaibrate a SPatial Econometric Interaction Model:
1.) A Spatial Weight
2.) A tibble data.frame cosnisiting of origins, destination, flows and distances between origins and destination
3.) Tiblee data.frame consisiting of the explanatory variables

### Retrieving data

```{r}
mpsz_nb <- read_rds("data/rds/mpsz_nb.rds")
mpsz_flow <- read_rds("data/rds/mpsz_flow.rds")
mpsz_var <- read_rds("data/rds/mpsz_var.rds")
```

### Creating spflow network class objects

```{r}
mpsz_net <- spflow_network(
  id_net ="sg",
  node_neighborhood = 
    nb2mat(mpsz_nb$by_contiguity),
  node_data = mpsz_var,
  node_key_column ="SZ_CODE")
mpsz_net
```


```{r}
mpsz_net_pairs <- spflow_network_pair(
  id_orig_net = "sg",
  id_dest_net ="sg",
  pair_data = mpsz_flow,
  orig_key_column = "ORIGIN_SZ",
  dest_key_column = "DESTIN_SZ"
)

mpsz_net_pairs

```

```{r}
mpsz_multi_net <- spflow_network_multi(mpsz_net, mpsz_net_pairs)
mpsz_multi_net
```
## Correlation Analysis
Multicollinearity checks for a situation which 2 more explanatory variavles are highly related in a regression model
```{r}
cor_formula <- log(1 + TRIPS) ~
  BUSSTOP_COUNT +
  AGE7_12 +
  AGE13_24 +
  AGE25_64 +
  SCHOOL_COUNT +
  BUSINESS_COUNT +
  RETAILS_COUNT +
  FINSERV_COUNT +
  P_(log(DISTANCE + 1))
  # Impedence from one point to another point
```

### Model calibration

```{r}
base_model <- spflow(
  spflow_formula = log(1 + TRIPS) ~
    O_(BUSSTOP_COUNT +
         AGE25_64) +
    D_(SCHOOL_COUNT +
         BUSINESS_COUNT +
         RETAILS_COUNT + 
         FINSERV_COUNT) +
    P_(log(DISTANCE + 1)),
  spflow_networks = mpsz_multi_net)
  
base_model

```
## Residual Diagnostic Tests
```{r}
old_par <- par(mfrow = c(1,3),
mar = c(2,2,2,2))
spflow_moran_plots(base_model)
```
```{r}
corr_residual <- pair_cor(base_model)
colnames(corr_residual) <- substr(colnames(corr_residual),1,3)
cor_image(corr_residual)
```
### Working with model control
```{r}
spflow_formula <- log(1 + TRIPS) ~
    O_(BUSSTOP_COUNT +
         AGE25_64) +
    D_(SCHOOL_COUNT +
         BUSINESS_COUNT +
         RETAILS_COUNT + 
         FINSERV_COUNT) +
    P_(log(DISTANCE + 1))

model_control <- spflow_control(
  estimation_method ="mle",
  model = "model_1")

mle_model1 <- spflow(
  spflow_formula,
  spflow_network = mpsz_multi_net,
  estimation_control = model_control)

mle_model1
```

```{r}
spflow_formula <- log(1 + TRIPS) ~
    O_(BUSSTOP_COUNT +
         AGE25_64) +
    D_(SCHOOL_COUNT +
         BUSINESS_COUNT +
         RETAILS_COUNT + 
         FINSERV_COUNT) +
    P_(log(DISTANCE + 1))

model_control <- spflow_control(
  estimation_method ="mle",
  model = "model_2")

mle_model2 <- spflow(
  spflow_formula,
  spflow_network = mpsz_multi_net,
  estimation_control = model_control)

mle_model2
```

```{r}
spflow_formula <- log(1 + TRIPS) ~
    O_(BUSSTOP_COUNT +
         AGE25_64) +
    D_(SCHOOL_COUNT +
         BUSINESS_COUNT +
         RETAILS_COUNT + 
         FINSERV_COUNT) +
    P_(log(DISTANCE + 1))

model_control <- spflow_control(
  estimation_method ="mle",
  model = "model_8")

mle_model8 <- spflow(
  spflow_formula,
  spflow_network = mpsz_multi_net,
  estimation_control = model_control)

mle_model8
```

```{r}
spflow_formula <- log(1 + TRIPS) ~
    O_(BUSSTOP_COUNT +
         AGE25_64) +
    D_(SCHOOL_COUNT +
         BUSINESS_COUNT +
         RETAILS_COUNT + 
         FINSERV_COUNT) +
    P_(log(DISTANCE + 1))

model_control <- spflow_control(
  estimation_method ="mle",
  model = "model_8")

mle_model9 <- spflow(
  spflow_formula,
  spflow_network = mpsz_multi_net,
  estimation_control = model_control)

mle_model9
```